% ============================================================
% ACKNOWLEDGEMENT
% ============================================================

\newpage
\chapter*{Acknowledgement}
\addcontentsline{toc}{chapter}{Acknowledgement}

Achievements, whether scholastic or otherwise, are not solely the result of individual effort but are shaped by the guidance, encouragement, and cooperation of many individuals. Several people, in their respective capacities, have contributed to the successful completion of this work.

We express our sincere gratitude to the Department of Computer Science and Engineering, RV College of Engineering, Bengaluru, for providing us the opportunity to undertake this project as part of the Network Programming and Security Lab course. This project enabled us to explore applied cryptography, secure system design, and network security concepts in a practical, real-world application.

We are grateful to Dr. K. N. Subramanya, Principal, RVCE, for his appreciation and support of this work. We extend our heartfelt thanks to Dr. Ramakanth Kumar P, Professor and Dean of Computer Science and Engineering, RVCE, and Dr. Shanta Rangaswamy, Head of the Department of Computer Science and Engineering, for their constant encouragement and support.

We express our deep sense of gratitude to our course instructor and project guide, Prof. Savitri Kulkarni, for her continuous guidance, valuable feedback, and encouragement throughout the development of this project. Her insights into network security and cryptography helped us refine our approach and maintain clarity in both the design and implementation.

We also thank all the teaching and technical staff of the Department of Computer Science and Engineering, RVCE, for their valuable assistance. Finally, we extend our appreciation to our teammates for their cooperation, discussions, and collaborative efforts, and to the institution for providing the necessary infrastructure and a conducive learning environment that made this work possible.
