% ============================================================
% BIBLIOGRAPHY
% ============================================================
\renewcommand{\bibname}{References}
\begin{thebibliography}{99}
\addcontentsline{toc}{chapter}{References}

\bibitem{webcrypto}
World Wide Web Consortium, ``Web Cryptography API,'' W3C Recommendation, 2017. [Online]. Available: \url{https://www.w3.org/TR/WebCryptoAPI/}

\bibitem{nist80038d}
National Institute of Standards and Technology, ``Recommendation for Block Cipher Modes of Operation: Galois/Counter Mode (GCM) and GMAC,'' NIST Special Publication 800-38D, 2007.

\bibitem{nist800132}
National Institute of Standards and Technology, ``Recommendation for Password-Based Key Derivation Part 1: Storage Applications,'' NIST Special Publication 800-132, 2010.

\bibitem{nyt}
A. Menezes, P. van Oorschot, and S. Vanstone, \textit{Handbook of Applied Cryptography}. CRC Press, 1996.

\bibitem{bernstein}
D. J. Bernstein, ``Curve25519: New Diffie-Hellman Speed Records,'' in \textit{Public Key Cryptography -- PKC 2006}, Springer, 2006, pp. 207--228.

\bibitem{rescorla}
E. Rescorla, ``Diffie-Hellman Key Agreement Method,'' IETF RFC 2631, 1999.

\bibitem{krawczyk}
H. Krawczyk, ``Cryptographic Extraction and Key Derivation: The HKDF Scheme,'' in \textit{Advances in Cryptology -- CRYPTO 2010}, Springer, 2010, pp. 631--648.

\bibitem{bellare}
M. Bellare and P. Rogaway, ``Introduction to Modern Cryptography,'' Lecture Notes, UCSD, 2005.

\bibitem{stallings}
W. Stallings, \textit{Cryptography and Network Security: Principles and Practice}, 8th ed. Pearson, 2020.

\bibitem{supabase}
Supabase, ``Supabase Documentation: Database, Auth, and Storage,'' 2024. [Online]. Available: \url{https://supabase.com/docs}

\bibitem{nextjs}
Vercel, ``Next.js Documentation: App Router,'' 2024. [Online]. Available: \url{https://nextjs.org/docs}

\bibitem{zustand}
P. Abdoul, ``Zustand: Bearable State Management for React,'' 2024. [Online]. Available: \url{https://github.com/pmndrs/zustand}

\end{thebibliography}
